% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      
I would like to express my gratitude to my supervisor Dr. Iasonas Triantis for his continuous support, guidance, knowledge, expert advice, patience and encouragement throughout the period   of my research.

I would also convey  my thanks and appreciation to my second supervisor Prof. Panayiotis Kyriacou, for sharing his knowledge and experience to complete the experimental procedures  to get  significant results.

I would like to acknowledge The School of Mathematics, Computer Science and Engineering at City, The University of London for financial assistance for this research work. . Thanks to its studentship I was able to complete this research, but it is also important to recognise the significant amount  of money that the school invested in the new laboratories and equipment which additionally helped me to complete my research  

Special appreciation  to my family, beginning  with my wife Nikki, my father Mario Enrique, my mother Eulalia, my sister Jennifer and my parents-in-law Terry and Bine. Thank you for their love, help, support and encouragement to complete this work. I am indebted to my parents for all their sacrifices made to secure my education.

Thanks to my friends and colleagues in the Research Centre for Biomedical Engineering for all their advice and sharing and learning knowledge amid  this time. Your remarks, comments and feedback were entirely helpful to complete this work. Special gratefulness  to my friend Tomas who was always there, to listen, to assist and give  any advice.

\begin{flushright}
	\textbf{Mario Guiovanni Bejarano Monroy}
\end{flushright}


\end{acknowledgements}
